\ \\ \LARGE\bf Speculative Evaluation in Particle Swarm Optimization \thanks{Matthew Gardner, Andrew McNabb, and Kevin Seppi are with the Department of Computer Science, Brigham Young University, 3361 TMCB, Provo, UT 84602 (phone: 801-422-8717; email: \{mjg82,a,k\}@cs.byu.edu).}

### 2010-03-18 honors_thesis/journal_paper/paper.tex
the commit that f9f984298 ('start alg') should have been

%SpecExPSO Stuff
\providecommand{\indic}{\ensuremath{I}}
\providecommand{\specvel}{\ensuremath{V}}
\providecommand{\specpos}{\ensuremath{X}}
\providecommand{\specval}{\ensuremath{FX}}
\providecommand{\ln}{\ensuremath{ln}}
\providecommand{\rn}{\ensuremath{rn}}
\providecommand{\caseset}{\ensuremath{\mathcal{C}}}
\providecommand{\casedef}{\ensuremath{(\pbest:\nbest)}}
\providecommand{\casexn}{\ensuremath{(x:\neigh*)}}
\providecommand{\casexx}{\ensuremath{(x:x)}}
\providecommand{\casexl}{\ensuremath{(x:\ln)}}
\providecommand{\casexr}{\ensuremath{(x:\rn)}}
\providecommand{\casepn}{\ensuremath{(\pers*:\neigh*)}}
\providecommand{\casepl}{\ensuremath{(\pers*:\ln)}}
\providecommand{\casepr}{\ensuremath{(\pers*:\rn)}}
\providecommand{\noeval}[1]{\ensuremath{#1^{-e}}}
\providecommand{\nonbest}[1]{\ensuremath{#1^{-n}}}
\providecommand{\p}{\ensuremath{p}}
\providecommand{\pset}{\ensuremath{\mathbf{p}}}
\providecommand{\s}{\ensuremath{s}}
\providecommand{\sset}{\ensuremath{\mathbf{s}}}
\providecommand{\nsset}{\ensuremath{\mathbf{ns}}}
\providecommand{\n}{\ensuremath{n}}
\providecommand{\nset}{\ensuremath{\mathbf{n}}}
\providecommand{\nnset}{\ensuremath{\mathbf{nn}}}

@@

Lastly, we also define notation for the value of the objective function at these points,
once again using $\casexn$ as an example:
\begin{align}
\label{eq:deffcasexn}
	\specval_{i,t+1}^{\casexn} & (\ppos_{i,t+1} ,\ppos_{\ln(i),t+1},\ppos_{\rn(i),t+1} ,\pbest_{i,t} ,\nbest_{i,t})= \\
\nonumber
	\specval_{i,t+1}^{\casexn} & =  \ofunc(\specval_{i,t+1}^{\casexn}
\end{align}

@@

The value of the objective function at $\ppos_{i,t+1}$ is given by:
\begin{align}
\label{eq:val2update}
	\ofunc (\ppos_{i,t+1}) = \sum_{c \in \caseset} \ \specval_{i,t+1}^{c} \ \indic_{i,t+1}^{c}
\end{align}
or, with $\specpos$ and $\indic$ written with the parameters which were omitted above:
\begin{align}
\nonumber
\ofunc & (\ppos_{i,t+1}) = \sum_{c \in \caseset} \\
\nonumber
& \ofunc(\specpos_{i,t+1}^{c}(\ppos_{i,t+1},\ppos_{\ln(i),t+1},\ppos_{\rn(i),t+1},
\pbest_{i,t},\nbest_{i,t},\pURand_{i,t}, \nURand_{i,t})) \\
\label{eq:val2updatelong}
& \indic_{i,t}^{c}(\ofunc ( \ppos_{i,t+1} ) ,\ofunc(\ppos_{\ln(i),t+1}),\ofunc(\ppos_{\rn(i),t+1}) ,\ofunc(\pbest_{i,t}) ,\ofunc(\nbest_{i,t}))
\end{align}

@@

\begin{algorithm} [h!]
  \caption{Standard PSO (minimizing)}
  \label{alg:standard}
  \begin{algorithmic}[1]
    \STATE t=0
    \STATE Initialize the swarm including
    \FOR{$i \in \swarm$}
      \STATE {$\ppos_{i,t} \leftarrow Random position$}
      \STATE {$\pvel_{i,t} \leftarrow Random velocities$}
      \STATE {$\fpbest_{i,t} \leftarrow Max Float$}
      \STATE {$\fnbest_{i,t} \leftarrow Max Float$}
    \ENDFOR
    \REPEAT
        \STATE $t=t+1$
        \FOR{$i \in \swarm$}
          \STATE {$\fppos_{i.t} \leftarrow \ofunc(\ppos_{i,t})$}
        \ENDFOR
        \FOR{$i \in \swarm$}
          \STATE $\nURand_{i,t} \Leftarrow RandomVector()$
          \STATE $\nURand_{i,t} \Leftarrow RandomVector()$
          \label{alg1:startifs}
          \STATE $\pbest_{i,t}\leftarrow \pbest_{i,t-1}$
          \STATE $\fpbest_{i,t}\leftarrow \fpbest_{i,t-1}$
          \IF {$\fppos_{i,t}<\fpbest_{i,t-1}$}
            \STATE $\pbest_{i,t}\leftarrow \ppos_{i,t}$
            \STATE $\fpbest_{i,t}\leftarrow \fppos_{i,t}$
          \ENDIF
          \STATE $\nbest_{i,t}\leftarrow \nbest_{i,t-1}$
          \STATE $\fnbest_{i,t}\leftarrow \fnbest_{i,t-1}$
          \IF {$\fppos_{i,t}<\fnbest_{i,t}$}
            \STATE $\pbest_{i,t}\leftarrow \ppos_{i,t}$
            \STATE $\fpbest_{i,t}\leftarrow \fppos_{i,t}$
          \ENDIF
          \IF {$\fppos_{\ln(i),t}<\fnbest_{i,t}$}
            \STATE $\pbest_{i,t}\leftarrow \ppos_{\ln(i),t}$
            \STATE $\fpbest_{i,t}\leftarrow \fppos_{\ln(i),t}$
          \ENDIF
          \IF {$\ofunc(\ppos_{i,t})<\ofunc(\nbest_{\rn(i),t})$}
            \STATE $\pbest_{i,t}\leftarrow \ppos_{\rn(i),t}$
            \STATE $\fpbest_{i,t}\leftarrow \fppos_{\rn(i),t}$
          \label{alg1:endifs}
          \ENDIF
          \STATE Update $\pvel_{i,t+1}$ as shown in equation~\eqref{eq:velupdate} 
          \STATE Update $\ppos_{i,t+1}$ as shown in equation~\eqref{eq:posupdate}
        \ENDFOR
    \UNTIL {$t >$ max iteration or convergence criteria met}
  \end{algorithmic}
\end{algorithm}

This algorithm can be transformed, will retaining it exact behavior, by
``unrolling'' one iteration of the ``Repeat'' loop, re-writing the ``if'''s in lines
\ref{alg1:startifs} through
\ref{alg1:endifs} to use the indicator functions $\indic_{i,t+1}$, and

\begin{algorithm} [h!]
  \caption{Unrolled PSO}
  \label{alg:unrolled}
  \begin{algorithmic}[1]
    \STATE t=0
    \STATE Initialize the swarm including
    \FOR{$i \in \swarm$}
      \STATE {$\ppos_{i,t} \leftarrow Random position$}
      \STATE {$\pvel_{i,t} \leftarrow Random velocities$}
      \STATE {$\fpbest_{i,t} \leftarrow Max Float$}
      \STATE {$\fnbest_{i,t} \leftarrow Max Float$}
    \ENDFOR
    \REPEAT
        \STATE $t=t+1$
        \FOR{$i \in \swarm$}
          \label{alg2:fatitert}
          \STATE DISPATCH {$\fppos_{i.t} \leftarrow \ofunc(\ppos_{i,t})$} to a seperate processor
        \ENDFOR
        \FOR{$i \in \swarm$}
          \STATE $\nURand_{i,t} \Leftarrow RandomVector()$
          \STATE $\nURand_{i,t} \Leftarrow RandomVector()$
          \FOR{$c \in \caseset$}
            \STATE Compute $\specvel_{i,t+1}^c$ and $\specpos_{i,t+1}^c$.
	    Examples shown in equations~\eqref{eq:defvcasexn} and~\eqref{eq:defpcasexn}
            \label{alg2:fatitertp1}
            \STATE DISPATCH computation of $\specval_{i,t+1}^c$ to a seperate processor.
	    Example shown in equation~\eqref{eq:deffcasexn}.
	    \COMMENT Note that none of the computations, $\specvel_{i,t+1}^c$,
	    $\specpos_{i,t+1}^c$ and $\specval_{i,t+1}^c$ depend on $\fppos_{i.t}$.
          \ENDFOR
        \ENDFOR
        \STATE WAIT for all function evaluations dispatched above to complete
        \FOR{$i \in \swarm$}
          \FOR{$c \in \caseset$}
            \STATE Compute $\indic_{i,t+1}^c$
	           Examples shown in equation~\eqref{eq:deficasexn} and~\eqref{eq:defpcasexn}
	  \ENDFOR 
          \STATE Compute $\pvel_{i,t+1}$ as shown in equation~\eqref{eq:vel2update} 
          \STATE Compute $\ppos_{i,t+1}$ as shown in equation~\eqref{eq:pos2update}
          \STATE Compute $\fppos_{i,t+1}$ as shown in equation~\eqref{eq:val2update}
        \ENDFOR

        \STATE $t=t+1$
        \FOR{$i \in \swarm$}
          \STATE $\nURand_{i,t} \Leftarrow RandomVector()$
          \STATE $\nURand_{i,t} \Leftarrow RandomVector()$
          \STATE $\pbest_{i,t}\leftarrow \pbest_{i,t-1}$
          \STATE $\fpbest_{i,t}\leftarrow \fpbest_{i,t-1}$
          \IF {$\fppos_{i,t}<\fpbest_{i,t-1}$}
            \STATE $\pbest_{i,t}\leftarrow \ppos_{i,t}$
            \STATE $\fpbest_{i,t}\leftarrow \fppos_{i,t}$
          \ENDIF
          \STATE $\nbest_{i,t}\leftarrow \nbest_{i,t-1}$
          \STATE $\fnbest_{i,t}\leftarrow \fnbest_{i,t-1}$
          \IF {$\fppos_{i,t}<\fnbest_{i,t}$}
            \STATE $\pbest_{i,t}\leftarrow \ppos_{i,t}$
            \STATE $\fpbest_{i,t}\leftarrow \fppos_{i,t}$
          \ENDIF
          \IF {$\fppos_{\ln(i),t}<\fnbest_{i,t}$}
            \STATE $\pbest_{i,t}\leftarrow \ppos_{\ln(i),t}$
            \STATE $\fpbest_{i,t}\leftarrow \fppos_{\ln(i),t}$
          \ENDIF
          \IF {$\ofunc(\ppos_{i,t})<\ofunc(\nbest_{\rn(i),t})$}
            \STATE $\pbest_{i,t}\leftarrow \ppos_{\rn(i),t}$
            \STATE $\fpbest_{i,t}\leftarrow \fppos_{\rn(i),t}$
          \ENDIF
          \STATE Update $\pvel_{i,t+1}$ as shown in equation~\eqref{eq:velupdate} 
          \STATE Update $\ppos_{i,t+1}$ as shown in equation~\eqref{eq:posupdate}
        \ENDFOR
    \UNTIL {$t >$ max iteration or convergence criteria met}
  \end{algorithmic}
\end{algorithm}

>>>>

%SpecExPSO Stuff
\providecommand{\indic}{\ensuremath{I}}
\providecommand{\specvel}{\ensuremath{V}}
\providecommand{\specpos}{\ensuremath{X}}
\providecommand{\ln}{\ensuremath{ln}}
\providecommand{\rn}{\ensuremath{rn}}
\providecommand{\caseset}{\ensuremath{\mathcal{C}}}
\providecommand{\casedef}{\ensuremath{(\pbest:\nbest)}}
\providecommand{\casexn}{\ensuremath{(x:\neigh*)}}
\providecommand{\casexx}{\ensuremath{(x:x)}}
\providecommand{\casexl}{\ensuremath{(x:\ln)}}
\providecommand{\casexr}{\ensuremath{(x:\rn)}}
\providecommand{\casepn}{\ensuremath{(\pers*:\neigh*)}}
\providecommand{\casepl}{\ensuremath{(\pers*:\ln)}}
\providecommand{\casepr}{\ensuremath{(\pers*:\rn)}}
\providecommand{\noeval}[1]{\ensuremath{#1^{-e}}}
\providecommand{\nonbest}[1]{\ensuremath{#1^{-n}}}
\providecommand{\p}{\ensuremath{p}}
\providecommand{\pset}{\ensuremath{\mathbf{p}}}
\providecommand{\s}{\ensuremath{s}}
\providecommand{\sset}{\ensuremath{\mathbf{s}}}
\providecommand{\nsset}{\ensuremath{\mathbf{ns}}}
\providecommand{\n}{\ensuremath{n}}
\providecommand{\nset}{\ensuremath{\mathbf{n}}}
\providecommand{\nnset}{\ensuremath{\mathbf{nn}}}

@@

@@

The value of the objective function at $\ppos_{i,t+1}$ is given by:
\begin{align}
\label{eq:pos2value}
	\ofunc (\ppos_{i,t+1}) = \sum_{c \in \caseset} \ \ofunc(\specpos_{i,t+1}^{c}) \ \indic_{i,t+1}^{c}
\end{align}
or, with $\specpos$ and $\indic$ written with the parameters which were omitted above:
\begin{align}
\nonumber
\ofunc & (\ppos_{i,t+1}) = \sum_{c \in \caseset} \\
\nonumber
& \ofunc(\specpos_{i,t+1}^{c}(\ppos_{i,t+1},\ppos_{\ln(i),t+1},\ppos_{\rn(i),t+1},
\pbest_{i,t},\nbest_{i,t},\pURand_{i,t}, \nURand_{i,t})) \\
\label{eq:pos2valuelong}
& \indic_{i,t}^{c}(\ofunc ( \ppos_{i,t+1} ) ,\ofunc(\ppos_{\ln(i),t+1}),\ofunc(\ppos_{\rn(i),t+1}) ,\ofunc(\pbest_{i,t}) ,\ofunc(\nbest_{i,t}))
\end{align}

@@

\begin{algorithm} [h!]
  \caption{Standard PSO (minimizing)}
  \label{alg:standard}
  \begin{algorithmic}[1]
    \STATE t=0
    \STATE Initialize the swarm including
    \FOR{$i \in \swarm$}
      \STATE {$\ppos_{i,t} \leftarrow Random position$}
      \STATE {$\pvel_{i,t} \leftarrow Random velocities$}
      \STATE {$\fpbest_{i,t} \leftarrow Max Float$}
      \STATE {$\fnbest_{i,t} \leftarrow Max Float$}
    \ENDFOR
    \REPEAT
        \STATE $t=t+1$
        \FOR{$i \in \swarm$}
          \STATE {$\fppos_{i.t} \leftarrow \ofunc(\ppos_{i,t})$}
        \ENDFOR
        \FOR{$i \in \swarm$}
          \label{alg1:startifs}
          \STATE $\pbest_{i,t}\leftarrow \pbest_{i,t-1}$
          \STATE $\fpbest_{i,t}\leftarrow \fpbest_{i,t-1}$
          \IF {$\fppos_{i,t}<\fpbest_{i,t-1}$}
            \STATE $\pbest_{i,t}\leftarrow \ppos_{i,t}$
            \STATE $\fpbest_{i,t}\leftarrow \fppos_{i,t}$
          \ENDIF
          \STATE $\nbest_{i,t}\leftarrow \nbest_{i,t-1}$
          \STATE $\fnbest_{i,t}\leftarrow \fnbest_{i,t-1}$
          \IF {$\fppos_{i,t}<\fnbest_{i,t}$}
            \STATE $\pbest_{i,t}\leftarrow \ppos_{i,t}$
            \STATE $\fpbest_{i,t}\leftarrow \fppos_{i,t}$
          \ENDIF
          \IF {$\fppos_{\ln(i),t}<\fnbest_{i,t}$}
            \STATE $\pbest_{i,t}\leftarrow \ppos_{\ln(i),t}$
            \STATE $\fpbest_{i,t}\leftarrow \fppos_{\ln(i),t}$
          \ENDIF
          \IF {$\ofunc(\ppos_{i,t})<\ofunc(\nbest_{\rn(i),t})$}
            \STATE $\pbest_{i,t}\leftarrow \ppos_{\rn(i),t}$
            \STATE $\fpbest_{i,t}\leftarrow \fppos_{\rn(i),t}$
          \ENDIF
          \STATE $\nURand_{i,t} \Leftarrow RandomVector()$
          \STATE $\nURand_{i,t} \Leftarrow RandomVector()$
          \STATE Update $\pvel_{i,t+1}$ as shown in equation~\eqref{eq:velupdate} 
          \STATE Update $\ppos_{i,t+1}$ as shown in equation~\eqref{eq:posupdate}
        \ENDFOR
    \UNTIL {$t >$ max iteration or convergence criteria met}
  \end{algorithmic}
\end{algorithm}

stuff \ref{alg1:startifs} more
